---
title: "Nicolai_ex03"
author: "Andrea Nicolai"
date: "20/4/2020"
output:
  pdf_document: default
  html_document: default
---

## Exercise 1

The time it takes a student to complete a TOLC-I University orientation and evaluation test follows a density function of the form:
$$
f (t) =
\begin{cases} 
c(t − 1)(2 − t) \qquad for\ 1 <t< 2 \\
0 \qquad \qquad \qquad \qquad otherwise
\end{cases}
$$
where t is the time in hours.

a) using the integrate() R function, determine the constant c (and verify it analytically)
```{r}
funct <- function(t) {
  temp <- ifelse ( t<2 & t>1 , 
                  (t-1)*(2-t),
                   0 )
  return(temp)
  }

funct.integral <- integrate(funct, lower = 1, upper = 2)
message(sprintf("The constant of integration c is: %.1f", 1/funct.integral$value))
```
Analitically we have that:

$$ \int_1^2 (t − 1)(2 − t)dt = 1/c  $$
That is 
$$ \left[ \frac{3t^2}{2} -2t - \frac{t^3}{3} \right]_1^2 = 1/c \quad \Rightarrow 1/6 = 1/c \Rightarrow \quad c = 6   $$ 
b) write the set of four R functions and plot the pdf and cdf, respectively

The cumulative computed analytically is:
$$ F(t) = 
\begin{cases}
 0 \qquad  \qquad  \qquad  \qquad  \qquad  \qquad t < 1 \\
 -2t^3 + 9t^2 -12t + 5  \qquad \quad 1 \leqslant t \leqslant 2\\
 1 \qquad  \qquad  \qquad  \qquad  \qquad  \qquad t > 2 \\
\end{cases} $$

```{r}
dprob1 <- function(t) { 
  temp <- ifelse ( t<2 & t>1, ((t-1)*(2-t)/func.integral$value), 0 )
  return(temp)
  }

pprob1 <- function(t) {
    tempcdf <- ifelse ( t<1, 0, ifelse(t<2 , -2*t^3 + 9*t^2 - 12*t + 5, 1) )
  return(tempcdf) 
}

rprob1 <- function(n) {
  u.1 <- runif(n, 1, 2)
  u.2 <- runif(n)
  func.max <- max(dprob1(seq(1,2,0.01)))
  y <- ifelse(u.2*func.max < dprob1(u.1), u.1, NA)
  y.clean <- y[!is.na(y)]
  return(y.clean)
}

  qprob1 <- function(q, n=1000) {
    u.1 <- runif(n, 1, 2)
    y <- ifelse(pprob1(u.1) < q, u.1, NA)
    y.clean <- y[!is.na(y)]
    return(max(y.clean))
  }
```

c) evaluate the probability that the student will finish the aptitude test in more than 75 minutes. And that it will take 90 and 120 minutes.
```{r}
message(sprintf("Probability that the student will finish the aptitude test in more than 75 minutes is: %.2f", 1-pprob1(1.25)))

message(sprintf("Probability that it will take 90 and 120 minutes is:  %.2f", pprob1(2)-pprob1(1.5)))
```


## Exercise 2

The lifetime of tires sold by an used tires shop is 10 4 · x km, where x is a random variable following the distribution function:
$$
f (x) =
\begin{cases}
\frac{1}{x^2} \ \qquad 1<x<2\\
0 \qquad \quad otherwise
\end{cases}
$$
Anlytically the CDF is :
$$
F(x) =
\begin{cases} 
0 \qquad\qquad\qquad\qquad\quad x < 1\\
2 \left(1 -\frac{2}{x^2}\right) \qquad for\ 1 \leqslant x \leqslant 2 \\
1 \qquad \qquad\qquad\qquad otherwise
\end{cases}
$$

Write the set of four R functions and plot the pdf and cdf, respectively.
```{r}
dprob2 <- function(x) {
  temp2 <- ifelse ( x<=2 & x>=1 , 2/(x^2), 0 )
  return(temp2)
  }

pprob2 <- function(x) {
  temp2cdf <- ifelse ( x<=1, 0, ifelse(x<2 , 2*(1-1/x) , 1) )
  return(temp2cdf) 
  }

rprob2 <- function(n) {
  us <- runif(n)
  temp2.r <- (2/(2-us))
  return(temp2.r)
  } 

qprob2 <- function(p) {
  temp2.q <- (2/(2-p))
  return(temp2.q)
}


```
```{r}
x <- seq(1,2, 0.01)

plot(x, dprob2(x), main = "pdf and cdf of the just defined distribution", col = "red", ylim = c(0,2.5), ylab = "prob", type = 'l')
points(x, pprob2(x), col = "blue", type = 'l')
legend( x = "top", c("pdf", "cdf") , col = c("red", "blue"), lty=1,  inset = 0.05)

```

Determine the probability that tires will last less than 15000 km
```{r}
message(sprintf("Probability that tires will last less than 15000 km is: %.2f ", pprob2(1.5)))
```
Sample 3000 random variables from the distribution and determine the mean value and the variance, using the expression $$Var(X) = E[X^2] − E[X]^2$$
```{r}
sample2 <- rprob2(3000)
  
integral.2 <- integrate( function(x) { 2/(x^2)*x } , 1 , 2)
message(sprintf("Mean from theory is: %.2f", integral.2$value))

E.X <- mean(sample2)
message(sprintf("Mean from sampling is: %.2f", E.X))

E.X2 <- mean(sample2*sample2)
VarX <- E.X2-(E.X)*(E.X)
message(sprintf("Variance from sampling is: %.2f", sqrt(E.X2)))

```

## Exercise 3

Markov’s inequality represents an upper boud to probability distributions:
$$
P (X ≥ k) ≤ \frac{E[X]}{k} \quad for \ k>0
$$
Having we defined a function:
$$
G(k) = 1 − F (k) ≡ P (X ≥ k)
$$
Plot G(k) and the Markov’s upper bound for:
  + The exponential, Exp(λ = 1), distribution function
  + The uniform, U (3, 5), distribution function
  + The binomial, Bin(n = 1, p = 1/2), distribution function
  + A Poisson, Pois(λ = 1/2), distribution function
```{r}


curve((1 - pexp( x , rate = 1)), from = 0, to = 12, col = "firebrick", ylab = "G(x)", main = "G(x) for exponential", ylim = c(0,3))
curve( 1/x , from = 0, to = 12, add = TRUE, col = 'chartreuse4')
legend( x = "top", c("pdf", "Markov's bound") , col = c("firebrick", "chartreuse4"), lty=1,  inset = 0.05)

curve((1 - punif(x,3,5)), from = 3, to = 5, col = "firebrick", ylab = "G(x)", main = "G(x) for uniform(3,5)", ylim = c(0,1.5))
curve( 4/x , from = 3, to = 5, add = TRUE, col = 'chartreuse4')
legend( x = "topright", c("pdf", "Markov's bound") , col = c("firebrick", "chartreuse4"), lty=1,  inset = 0.05)

curve((1 - pbinom(x, size = 1, prob = 1/2)), from = 0, to = 12, col = "firebrick", ylab = "G(x)", main = "G(x) for binomial ( n=1, p= 0.5)", ylim = c(0,3))
curve( 0.5/x , from = 0, to = 12, add = TRUE, col = 'chartreuse4')
legend( x = "top", c("pdf", "Markov's bound") , col = c("firebrick", "chartreuse4"), lty=1,  inset = 0.05)

curve((1 - ppois( x , lambda = 1/2)), from = 0, to = 6, col = "firebrick", ylab = "G(x)", main = "G(x) for Poissonian (lambda=0.5)", ylim = c(0, 3))
curve( 1/(2*x) , from = 0, to = 6, add = TRUE, col = 'chartreuse4')
legend( x = "top", c("pdf", "Markov's bound") , col = c("firebrick", "chartreuse4"), lty=1,  inset = 0.05)
```

## Exercise 4

Chebyshev’s inequality tells us that:
$$
P ( |X − μ| ≥ kσ) ≤ \frac{1}{k^2}
$$
which can be rewritten as:
$$
P ( |X − μ| < kσ) ≥ 1 − \frac{1}{k^2}
$$
Use R to show, with a plot, that Chebyshev’s inequality is is an upper bound to the following distributions:

  + a normal distribution, N(μ = 3, σ = 5)
  + an exponential distribution, Exp(λ = 1)
  + a uniform distribution U (1 −√2, 1 + √2)
  + a Poisson, Pois(λ = 1/3), distribution function

```{r}

curve(pnorm((5*x+3),3,5) - pnorm((3-5*x),3,5), 0, 10 , main = "Normal (mean = 3, sigma = 5)", col = "forestgreen",ylab= "")
curve(1-1/x^2, 0, 10, add=TRUE, col = 'deeppink')
legend( x = "right", c("pdf", "Chebyshev's bound") , col = c("forestgreen", "deeppink"), lty=1,  inset = 0.05)

curve((pexp(1+x, rate = 1) - pexp( 1-x , rate = 1)), from = 0, to = 12, col = "forestgreen", main = "exponential (rate=1)",ylab= "")
curve( 1- 1/(x^2) , from = 0, to = 12, add = TRUE, col = 'deeppink')
legend( x = "right", c("pdf", "Chebyshev's bound") , col = c("firebrick", "deeppink"), lty=1,  inset = 0.05)

curve(punif( 2*x/3 + 1 , 1-sqrt(2), 1+sqrt(2)) - punif( -2*x/3 + 1 , 1-sqrt(2), 1+sqrt(2)), from = 0, to = 12, col = "forestgreen", main = "unif (1-sqrt2, 1+sqrt2)", ylab= "")
curve( 1- 1/(x^2) , from = 0, to = 12, add = TRUE, col = 'deeppink')
legend( x = "right", c("pdf", "Chebyshev's bound") , col = c("forestgreen", "deeppink"), lty=1,  inset = 0.05)

plot( (0:5) , ppois((0:5)*(1/3) + 1/3, 1/3) , col = "forestgreen", main = "poisson (rate=1/3)", ylab= "", type = 'b')
curve( 1- 1/(x^2) , from = 0, to = 6, add = TRUE, col = 'deeppink')
legend( x = "bottomright", c("pdf", "Chebyshev's bound") , col = c("forestgreen", "deeppink"), lty=1,  inset = 0.05)
```





